The ChEMBL database as Linked Open Data

### 2012-07-17 article.tex
Added Bin's contribution

David/Bin: Chem2Bio2RDF

>>>>

Chem2Bio2RDF created a single RDF repository covering over 20 public data resources pertaining to drugs, chemical compounds, carcinogens,
protein targets, genes, diseases, side effects, pathways and their relations~\cite{Chen2010}. The entities and their relations were
further annotated by Chem2Bio2OWL ontology, making it a rich semantic resource for integrative search~\cite{Chen2012a} and
mining~\cite{Chen2012b}.  ChEMBL-RDF set (version 13) has been uploaded into Chem2Bio2RDF triple store, enabling to perform
queries linking ChEMBL with other entities in Chem2Bio2RDF. Since both ChEMBL RDF and Chem2Bio2RDF use InChi keys to present chemical
compounds and adopt Bio2RDF protein identifiers to present targets, queries can be easily constructed to link each other. For instance,
to investigate the relations between drug side effects and their targets (usually off-targets), a query was created to link the
targets in ChEMBL to side effects (e.g., heart disease) in Chem2Bio2RDF via side effect related drugs and their bioassay activities.
In this case, 36 drugs causing heart diseases were linked to 87 unique protein targets (IC$_50$ $<$ 10um). The top two most common targets
of these drugs are HERG and Serotonin transporter. HERG is a well-known off-target accounting for cardiovascular problems, while
the activity of Serotonin transporter may also be associated with significant cardiovascular adverse effects~\cite{Levy2006}. 
The targets can be also linked to side effects via protein protein interactions by searching across ChEMBL, HPRD and Sider in
Chem2Bio2RDF. SPARQLs are available at \url{http://chem2bio2rdf.wikispaces.com/Integrate+with+ChEMBL} and can be executed in the
SPARQL endpoint~\cite{Chen2010}.
